---
title: "Exploratory Analysis for Georeferencing Project"
output:
  html_document:
    code_folding: hide
    df_print: kable
---
```{r message = FALSE}
# Load core libraries; install these packages if you have not already
library(ridigbio)
library(rgbif)
library(tidyverse)

# Load library for making nice HTML output
library(kableExtra)
```

### Big picture

```{r}
# Get count of mammal specimen records on iDigBio
mammalia_idb <- idig_count_records(rq = list(class = "mammalia"))

# Get count of mammal specimen records on GBIF
mammalia_gbif <- occ_count(taxonKey = 359,
                           basisOfRecord = "PRESERVED_SPECIMEN")

# Get count of mammal specimen records that are georeferenced on iDigBio
mammalia_georef_idb <- idig_count_records(rq = list(class = "mammalia",
                                                    geopoint = list(
                                                      type = "exists")))

# Get count of mammal specimen records that are georeferenced on GBIF
mammalia_georef_gbif <- occ_count(taxonKey = 359,
                                  basisOfRecord = "PRESERVED_SPECIMEN",
                                  georeferenced = TRUE)

# Get count of mammal specimens that are not georeferenced but do have values
# in the locality field on iDigBio
mammalia_locality_idb <- idig_count_records(rq = list(class = "mammalia",
                                                      geopoint = list(
                                                        type = "missing"),
                                                      locality = list(
                                                        type = "exists")))

# TBD: Get count of mammal specimens that are not georeferenced but do have values
# in the locality field on GBIF
mammalia_locality_gbif <- "TBD"
```

_**How many specimen records for Mammalia exist on iDigBio and GBIF?**_

- **iDigBio** = `r format(mammalia_idb, big.mark = ",")`
- **GBIF** = `r format(mammalia_gbif, big.mark = ",")`

_**How many of these specimen records are georeferenced already?**_

- **iDigBio** = `r format(mammalia_georef_idb, big.mark = ",")` (`r round(sum(mammalia_georef_idb/mammalia_idb*100), 1)`%)
- **GBIF** = `r format(mammalia_georef_gbif, big.mark = ",")` (`r round(sum(mammalia_georef_gbif/mammalia_gbif*100), 1)`%)

_**Of those specimen records that are not already georeferenced, how many have something in the locality field?**_

- **iDigBio** = `r format(mammalia_locality_idb, big.mark = ",")` (`r round(sum(mammalia_locality_idb/(mammalia_idb-mammalia_georef_idb)*100), 1)`%)
- **GBIF** = `r mammalia_locality_gbif`

### Scoped by taxonomic family of interest

**Primary taxonomic scope**: Rhinolophidae, Hipposideridae

**Secondary taxonomic scope**: Megadermatidae, Rhinopomatidae, Craseonycteridae

**Tertiary taxonomic scope**: Pteropodidae

**Broadest taxonomic scope**: TBD civet and pangolins (Pholidota) taxa

For each family, we want to know:

1. How many specimen records exist on iDigBio and on GBIF?
1. How many of these records are already georeferenced?
1. Of those records that are not already georeferenced, how many have something in the locality field?
1. How many distinct collections are providing these specimen records?
    - Who are the top five data providers based on total number of records?
    - What percentage of records are contributed by these top five?
    - Who are the top five data providers based on number of records needing to be georeferenced?
    - What percentage of records needing to be georeferenced are contributed by these top five?
1. Where are these recordsets being mobilized from? (e.g. the VertNet IPT)
1. What is the geographic scope represented in these records, measured by country of collection?
    - All records
    - Records in need of georeferencing
1. What is the temporal scope represented in these records, measured by year of collection?
    - All records
    - Records in need of georeferencing
1. What is the taxonomic scope represented in these records, measured by distinct species?
    - All records
    - Records in need of georeferencing
1. What kind of ancillary locality information is represented in these records, measured by presence of values in the habitat field?
    - All records
    - Records in need of georeferencing
1. What kind of ancillary locality information is represented in these records, measured by presence of values in the preparations and associatedSequences fields?
    - All records
    - Records in need of georeferencing

```{r}
queryByFamily <- function(family) {
  
  # Find GBIF taxonKey for given family
  taxonKey <- name_backbone(name = family,
                            rank = "family",
                            class = "mammalia")
  
  # Rename taxonKey dataframe with the input family value and make it available
  # in the global environment
  assign(paste("taxonKey", family, sep = "_"), taxonKey, envir = globalenv())
  
  # Get count of specimen records on iDigBio
  idb <- idig_count_records(rq = list(family = family))
  
  # Get count of specimen records that are georeferenced on iDigBio
  georef_idb <- idig_count_records(rq = list(family = family,
                                             geopoint = list(
                                               type = "exists")))
  
  # Get count of specimens that are not georeferenced but do have values
  # in the locality field on iDigBio
  locality_idb <- idig_count_records(rq = list(family = family,
                                               geopoint = list(
                                                 type = "missing"),
                                               locality = list(
                                                 type = "exists")))
  
  # Get count of specimen records on GBIF
  gbif <- occ_count(taxonKey = taxonKey$usageKey[1],
                    basisOfRecord = "PRESERVED_SPECIMEN")
  
  # Get count of specimen records that are georeferenced on GBIF
  georef_gbif <- occ_count(taxonKey = taxonKey$usageKey[1],
                           basisOfRecord = "PRESERVED_SPECIMEN",
                           georeferenced = TRUE)
  
  # Get count of specimens that are not georeferenced but do have values
  # in the locality field on GBIF
  locality_gbif <- "TBD"
  
  # Create table to present variables above
  summary <- tribble(
    ~source,
    ~totalCount,
    ~georefCount,
    ~georefPercent, 
    ~notGeorefCount,
    ~notGeorefPercent,
    ~couldGeorefCount,
    ~couldGeorefPercent,
    "iDigBio",
    idb,
    georef_idb,
    paste(round(sum(georef_idb/idb*100), 1), "%", sep = ""),
    sum(idb-georef_idb),
    paste(round(sum((idb-georef_idb)/idb*100), 1), "%", sep = ""),
    locality_idb,
    paste(round(sum(locality_idb/(idb-georef_idb)*100), 1), "%", sep = ""),
    "GBIF",
    gbif,
    georef_gbif,
    paste(round(sum(georef_gbif/gbif*100), 1), "%", sep = ""),
    sum(gbif-georef_gbif),
    paste(round(sum((gbif-georef_gbif)/gbif*100), 1), "%", sep = ""),
    locality_gbif,
    "TBD")
    
  # Rename records dataframe with the input family value and make it available
  # in the global environment
  assign(paste("summary", family, sep = "_"), summary, envir = globalenv())
  
  # Pull in specimen records from iDigBio that are not georeferenced but that do
  # have values in the locality field
  records_idb <- idig_search_records(rq = list(family = family,
                                               geopoint = list(
                                                 type = "missing"),
                                               locality = list(
                                                 type = "exists")),
                                     fields = c("uuid",
                                                "recordset",
                                                "institutioncode",
                                                "collectioncode",
                                                "catalognumber",
                                                "scientificname",
                                                "eventdate",
                                                "country",
                                                "stateprovince",
                                                "county",
                                                "municipality",
                                                "data.dwc:locality",
                                                "data.dwc:habitat",
                                                "data.dwc:preparations",
                                                "data.dwc:associatedSequences"))
  
  # Create column in records_idb to match duplicate records with GBIF data
  records_idb <- records_idb %>% 
    unite(matchDuplicates, 
          c("institutioncode", "collectioncode", "catalognumber"), 
          sep = " ", 
          remove = FALSE) %>% 
    mutate(matchDuplicates = tolower(matchDuplicates))
  
  # Pull in specimen records from GBIF that are not georeferenced
  records_gbif <- occ_search(taxonKey = taxonKey$usageKey[1],
                             basisOfRecord = "PRESERVED_SPECIMEN",
                             hasCoordinate = FALSE,
                             fields = c("key",
                                        "datasetKey",
                                        "institutionCode",
                                        "collectionCode",
                                        "catalogNumber",
                                        "acceptedScientificName",
                                        "year",
                                        "country",
                                        "stateProvince",
                                        "county",
                                        "municipality",
                                        "locality",
                                        "habitat",
                                        "preparations",
                                        "associatedSequences"))
  
  # Extract dataframe from records_gbif list
  records_gbif <- records_gbif$data
  
  # Subset records_gbif to include only those records with locality values,
  # and create column to match duplicate records with iDigBio data
  records_gbif <- records_gbif %>% 
    filter(!is.na(locality)) %>% 
    unite(matchDuplicates, 
          c("institutionCode", "collectionCode", "catalogNumber"), 
          sep = " ", 
          remove = FALSE) %>% 
    mutate(matchDuplicates = tolower(matchDuplicates))
  
  # Join records_idb and records_gbif
  records <- records_idb %>% 
    full_join(records_gbif, by = "matchDuplicates")
  
  # Rename records dataframe with the input family value and make it available
  # in the global environment
  assign(paste("records", family, sep = "_"), records, envir = globalenv())
  
}

queryByFamily("Rhinolophidae")
queryByFamily("Hipposideridae")
queryByFamily("Megadermatidae")
queryByFamily("Rhinopomatidae")
queryByFamily("Craseonycteridae")
queryByFamily("Pteropodidae")
```

#### RHINOLOPHIDAE

_**Summary of data on iDigBio and GBIF**_

```{r echo = FALSE, results = 'asis'}
knitr::kable(summary_Rhinolophidae) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"))
```

#### HIPPOSIDERIDAE

_**Summary of data on iDigBio and GBIF**_

```{r echo = FALSE, results = 'asis'}
knitr::kable(summary_Hipposideridae) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"))
```

#### MEGADERMATIDAE

_**Summary of data on iDigBio and GBIF**_

```{r echo = FALSE, results = 'asis'}
knitr::kable(summary_Megadermatidae) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"))
```

#### RHINOPOMATIDAE

_**Summary of data on iDigBio and GBIF**_

```{r echo = FALSE, results = 'asis'}
knitr::kable(summary_Rhinopomatidae) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"))
```

#### CRASEONYCTERIDAE

_**Summary of data on iDigBio and GBIF**_

```{r echo = FALSE, results = 'asis'}
knitr::kable(summary_Craseonycteridae) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"))
```

#### PTEROPODIDAE

_**Summary of data on iDigBio and GBIF**_

```{r echo = FALSE, results = 'asis'}
knitr::kable(summary_Pteropodidae) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"))
```
