---
title: "Exploratory Analysis for Georeferencing Project"
output:
  html_document:
    code_folding: hide
    df_print: kable
---

```{r message = FALSE}
# Load core libraries; install these packages if you have not already
library(ridigbio)
library(rgbif)
library(tidyverse)

# Load library for making nice HTML output
library(kableExtra)
```
```{r global_options, include = FALSE}
# Set global chunk options
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      results = 'asis')
```

----------

### Primary taxonomic scope

----------

RHINOLOPHIDAE, HIPPOSIDERIDAE, RHINONYCTERIDAE

```{r message = FALSE, warning = FALSE}
# Read into R the raw occurrence data from iDigBio, which should be whatever was
# published by the data provider (e.g. the collection)
idb_raw <- read_csv("77dfb340-4ab1-4c06-9f0c-eecc0558a9c8/occurrence_raw.csv", 
                    na = character(),
                    col_types = cols())

# Rename columns to reflect provenance and remove illegal characters
idb_raw <- idb_raw %>% 
  rename_all(function(x){paste0(x, "_idbR")}) %>% 
  rename_all(funs(str_replace_all(., "dwc:", ""))) %>% 
  rename_all(funs(str_replace_all(., ":", "_")))

# Read into R the version of occurrence data processed by iDigBio
idb_processed <- read_csv("77dfb340-4ab1-4c06-9f0c-eecc0558a9c8/occurrence.csv", 
                          na = character(),
                          col_types = cols())

# Rename columns to reflect provenance and remove illegal characters
idb_processed <- idb_processed %>% 
  rename_all(function(x){paste0(x, "_idbP")}) %>% 
  rename_all(funs(str_replace_all(., "dwc:", ""))) %>% 
  rename_all(funs(str_replace_all(., ":", "_"))) 

# Join raw and processed iDigBio data together
idb_joined <- idb_raw %>% 
  left_join(idb_processed, by = c("coreid_idbR" = "coreid_idbP"))

# Subset iDigBio data by families of interest
idb_subset <- idb_joined %>% 
  mutate(family_idbR = tolower(family_idbR)) %>% 
  filter(family_idbR %in% c("rhinolophidae", 
                             "hipposideridae", 
                             "rhinonycteridae") |
          family_idbP %in% c("rhinolophidae", 
                             "hipposideridae", 
                             "rhinonycteridae")) %>% 
  unite(matchDuplicates2, 
          c("institutionCode_idbR", 
            "collectionCode_idbR", 
            "catalogNumber_idbR"), 
          sep = " ", 
          remove = FALSE) %>% 
  mutate(matchDuplicates2 = str_squish(tolower(matchDuplicates2))) %>%
  group_by(matchDuplicates2) %>% 
  filter() %>% 
  mutate(matchDuplicates = case_when(row_number() != 1 
                                     ~ paste("PLACEHOLDER_IDB_", 
                                           matchDuplicates2,
                                           sep = ""),
                                     row_number() == 1 ~ matchDuplicates2)) %>% 
  ungroup() %>% 
  mutate(matchDuplicates = make.unique(matchDuplicates))

# Read into R occurrence data from GBIF
gbif_subset_1 <- read_tsv("0052900-200221144449610/occurrence.txt", 
                     na = character(),
                     col_types = cols())

# Rename columns to reflect provenance and remove illegal characters
gbif_subset_1 <- gbif_subset_1 %>% 
  rename_all(function(x){paste0(x, "_gbif")})

# Read into R occurrence data from GBIF
gbif_subset_2 <- read_tsv("0052902-200221144449610/occurrence.txt", 
                     na = character(),
                     col_types = cols())

# Rename columns to reflect provenance and remove illegal characters
gbif_subset_2 <- gbif_subset_2 %>% 
  rename_all(function(x){paste0(x, "_gbif")})

# Bind `gbif_1` and `gbif_2` together
gbif_subset <- bind_rows(list(gbif_subset_1, gbif_subset_2), .id = "id") %>% 
  unite(matchDuplicates2, 
          c("institutionCode_gbif", 
            "collectionCode_gbif", 
            "catalogNumber_gbif"), 
          sep = " ", 
          remove = FALSE) %>% 
  mutate(matchDuplicates2 = str_squish(tolower(matchDuplicates2))) %>%
  group_by(matchDuplicates2) %>% 
  filter() %>% 
  mutate(matchDuplicates = case_when(row_number() != 1 
                                     ~ paste("PLACEHOLDER_GBIF_", 
                                           matchDuplicates2,
                                           sep = ""),
                                     row_number() == 1 ~ matchDuplicates2)) %>%
  ungroup() %>% 
  mutate(matchDuplicates = make.unique(matchDuplicates))

# Join `idb_subset` and `gbif_subset` based on unique combinations of intitution
# code, collection code, and catalog number
records <- gbif_subset %>% 
  full_join(idb_subset, by = "matchDuplicates") %>% 
  mutate(idigbio = case_when(!is.na(idigbio_uuid_idbP) ~ "yes")) %>% 
  mutate(gbif = case_when(!is.na(gbifID_gbif) ~ "yes")) %>% 
  select(gbif, idigbio, gbifID_gbif, idigbio_uuid_idbP, 
         sort(tidyselect::peek_vars()),
         -starts_with("matchDuplicates"))

# Subset records that overlap between iDigBio and GBIF for visualizing data
records_overlap <- records %>% 
  filter(gbif == "yes" & idigbio == "yes")

# Count records that have already been georeferenced 
records_georef <- records %>% 
  filter(idigbio_geoPoint_idbP != "" | 
         (!is.na(decimalLatitude_gbif) & !is.na(decimalLongitude_gbif))) %>% 
  nrow()

# Count distinct localities that have already been georeferenced 
localities_georef <- records %>% 
  filter(idigbio_geoPoint_idbP != "" | 
         (!is.na(decimalLatitude_gbif) & !is.na(decimalLongitude_gbif))) %>% 
  unite(locality,
        c("locality_gbif", "locality_idbR", "locality_idbP"),
        sep = " | ") %>% 
  count(locality) %>% 
  arrange(desc(n))

# Subset records that could be georeferenced, as indicated by a lack of 
# existing georeference and a value in the locality field
records_locality <- records %>% 
  filter((idigbio_geoPoint_idbP == "" | is.na(idigbio_geoPoint_idbP)) & 
         (is.na(decimalLatitude_gbif) & is.na(decimalLongitude_gbif))) %>% 
  filter((!is.na(locality_gbif) & locality_gbif != "" ) |
         (!is.na(locality_idbR) & locality_idbR != "" ) |
         (!is.na(locality_idbP) & locality_idbP != "" ) |
         (!is.na(verbatimLocality_gbif) & verbatimLocality_gbif != "" ) |
         (!is.na(verbatimLocality_idbR) & verbatimLocality_idbR != ""))

# Count distinct localities that have not yet been georeferenced 
localities_locality <- records_locality %>% 
  unite(locality,
        c("locality_gbif", "locality_idbR", "locality_idbP"),
        sep = " | ") %>% 
  count(locality) %>% 
  arrange(desc(n))

# Count distinct data providers of total records
providers_total <- records %>% 
  mutate(institutionCode = coalesce(institutionCode_gbif, 
                                    institutionCode_idbR, 
                                    institutionCode_idbP)) %>% 
  count(institutionCode) %>% 
  arrange(desc(n)) %>% 
  rename(records = n) %>% 
  mutate(percentRecords = round(records/(sum(records))*100, 1))

# Count distinct data providers of ungeoreferenced records
providers_ungeoreferenced <- records_locality %>% 
  mutate(institutionCode = coalesce(institutionCode_gbif, 
                                    institutionCode_idbR, 
                                    institutionCode_idbP)) %>% 
  count(institutionCode) %>% 
  arrange(desc(n)) %>% 
  rename(records = n) %>% 
  mutate(percentRecords = round(records/(sum(records))*100, 1))

# Count distinct countries represented in total records
countries_total <- records %>% 
  mutate(country = coalesce(country_idbP,
                            countryCode_gbif,
                            country_idbR)) %>% 
  count(country) %>% 
  arrange(desc(n)) %>% 
  rename(records = n) %>% 
  mutate(percentRecords = round(records/(sum(records))*100, 1))
  
# Count distinct countries represented in ungeoreferenced records
countries_ungeoreferenced <- records_locality %>% 
  mutate(country = coalesce(country_idbP,
                            countryCode_gbif,
                            country_idbR)) %>% 
  count(country) %>% 
  arrange(desc(n)) %>% 
  rename(records = n) %>% 
  mutate(percentRecords = round(records/(sum(records))*100, 1))

# Count distinct species represented in total records
species_total <- records %>% 
  mutate(species = coalesce(scientificName_gbif,
                            scientificName_idbP,
                            scientificName_idbR)) %>% 
  mutate(species = tolower(species)) %>% 
  mutate(family = coalesce(family_gbif,
                           family_idbP,
                           family_idbR)) %>% 
  mutate(family = toupper(family)) %>% 
  group_by(family, species) %>% 
  count(species) %>% 
  rename(records = n) %>% 
  ungroup() %>% 
  mutate(percentRecords = round(records/(sum(records))*100, 1))
  
# Count distinct species represented in ungeoreferenced records
species_ungeoreferenced <- records_locality %>% 
  mutate(species = coalesce(scientificName_gbif,
                            scientificName_idbP,
                            scientificName_idbR)) %>% 
  mutate(species = tolower(species)) %>% 
  mutate(family = coalesce(family_gbif,
                           family_idbP,
                           family_idbR)) %>% 
  mutate(family = toupper(family)) %>% 
  group_by(family, species) %>% 
  count(species) %>% 
  rename(records = n) %>% 
  ungroup() %>% 
  mutate(percentRecords = round(records/(sum(records))*100, 1))

# Read into R data from batnames resource
batnames_hipposideridae <- read_csv("Hipposideridae2020-04-28.csv", 
                    na = character(),
                    col_types = cols())

# Combine batnames columns for matching
batnames_hipposideridae <- batnames_hipposideridae %>% 
  unite(match,
        c("Genus", "Species"),
        sep = " ",
        remove = FALSE) %>% 
  mutate(match = tolower(match)) %>% 
  unite(matchAuthor,
        c("Genus", "Species", "Author"),
        sep = " ",
        remove = FALSE) %>% 
  mutate(matchAuthor = tolower(matchAuthor)) %>% 
  unite(matchFullAuthor,
        c("matchAuthor", "Date"),
        sep = ", ",
        remove = FALSE) %>% 
  mutate(matchFullAuthor = tolower(matchFullAuthor))

# Read into R data from batnames resource
batnames_rhinonolophidae <- read_csv("Rhinolophidae2020-04-28.csv", 
                    na = character(),
                    col_types = cols())

# Combine batnames columns for matching
batnames_rhinonolophidae <- batnames_rhinonolophidae %>% 
  unite(match,
        c("Genus", "Species"),
        sep = " ",
        remove = FALSE) %>% 
  mutate(match = tolower(match)) %>% 
  unite(matchAuthor,
        c("Genus", "Species", "Author"),
        sep = " ",
        remove = FALSE) %>% 
  mutate(matchAuthor = tolower(matchAuthor)) %>% 
  unite(matchFullAuthor,
        c("matchAuthor", "Date"),
        sep = ", ",
        remove = FALSE) %>% 
  mutate(matchFullAuthor = tolower(matchFullAuthor))

# Read into R data from batnames resource
batnames_rhinonycterideridae <- read_csv("Rhinonycteridae2020-04-28.csv", 
                    na = character(),
                    col_types = cols())

# Combine batnames columns for matching
batnames_rhinonycterideridae <- batnames_rhinonycterideridae %>% 
  unite(match,
        c("Genus", "Species"),
        sep = " ",
        remove = FALSE) %>% 
  mutate(match = tolower(match)) %>% 
  unite(matchAuthor,
        c("Genus", "Species", "Author"),
        sep = " ",
        remove = FALSE) %>% 
  mutate(matchAuthor = tolower(matchAuthor)) %>% 
  unite(matchFullAuthor,
        c("matchAuthor", "Date"),
        sep = ", ",
        remove = FALSE) %>% 
  mutate(matchFullAuthor = tolower(matchFullAuthor))

# Bind all batnames datasets into a single data frame
batnames <- bind_rows(batnames_hipposideridae,
                      batnames_rhinonolophidae,
                      batnames_rhinonycterideridae)

# Check to see which species have matching names in Batnames data
species_total <- species_total %>% 
  left_join(select(batnames, Family, match), 
            by = c("species" = "match")) %>% 
  left_join(select(batnames, Family, matchAuthor), 
            by = c("species" = "matchAuthor")) %>% 
  left_join(select(batnames, Family, matchFullAuthor), 
            by = c("species" = "matchFullAuthor")) %>% 
  mutate(familyBatnames = coalesce(Family, Family.x, Family.y)) %>% 
  mutate(familyBatnames = toupper(familyBatnames)) %>% 
  select(family, familyBatnames, species, records, percentRecords)

# Check to see which species have matching names in Batnames data
species_ungeoreferenced <- species_ungeoreferenced %>% 
  left_join(select(batnames, Family, match), 
            by = c("species" = "match")) %>% 
  left_join(select(batnames, Family, matchAuthor), 
            by = c("species" = "matchAuthor")) %>% 
  left_join(select(batnames, Family, matchFullAuthor), 
            by = c("species" = "matchFullAuthor")) %>% 
  mutate(familyBatnames = coalesce(Family, Family.x, Family.y)) %>% 
  mutate(familyBatnames = toupper(familyBatnames)) %>% 
  select(family, familyBatnames, species, records, percentRecords)

# Calculate percent names valid according to Batnames
records_total_batnames <- species_total %>% 
  filter(!is.na(familyBatnames)) %>% 
  summarize(valid = sum(records)) %>% 
  as.integer()

# Calculate percent names valid according to Batnames
records_ungeoreferenced_batnames <- species_ungeoreferenced %>% 
  filter(!is.na(familyBatnames)) %>% 
  summarize(valid = sum(records)) %>% 
  as.integer()

# Count records that have a value in the habitat field
records_habitat <- records %>% 
  filter(habitat_gbif != "" | 
         habitat_idbR != "" ) %>% 
  nrow()

# Count records that have a value in the preparations field
records_prep <- records %>% 
  filter(preparations_gbif != "" | 
         preparations_idbR != "" ) %>% 
  nrow()

# Count records that have a value in the habitat field
records_seq <- records %>% 
  filter(associatedSequences_gbif != "" | 
         associatedSequences_idbR != "" ) %>% 
  nrow()
```
----------

_**How many specimen records exist?**_

----------

- iDigBio = `r format(nrow(idb_subset), big.mark = ",")`
- GBIF = `r format(nrow(gbif_subset), big.mark = ",")`
- Aggregators combined, accounting for obvious duplicates = `r format(nrow(records), big.mark = ",")`

_Example of records that exist on both iDigBio and GBIF and have been merged here_

```{r echo = FALSE}
knitr::kable(head(records_overlap)) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "200px")
```

&nbsp;

_`r nrow(as_tibble(colnames(records)))` fields present in the aggregator combined data_

```{r echo = FALSE}
as_tibble(colnames(records)) %>% 
knitr::kable() %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "50%", height = "400px")
```

&nbsp;

```{r echo = FALSE}
col <- as_tibble(colnames(records)) %>% 
  nrow()
col_gbif <- as_tibble(colnames(records)) %>% 
  filter(grepl("_gbif", value)) %>% 
  nrow()
col_idbR <- as_tibble(colnames(records)) %>% 
  filter(grepl("_idbR", value)) %>% 
  nrow()
col_idbP <- as_tibble(colnames(records)) %>% 
  filter(grepl("_idbP", value)) %>% 
  nrow()
```

_Origin of the fields present in the aggregator combined data:_

- `r round(sum(col_gbif/col*100), 1)`% from GBIF compiled data
- `r round(sum(col_idbR/col*100), 1)`% from iDigBio raw data
- `r round(sum(col_idbP/col*100), 1)`% from iDigBio processed

----------

_**How many of these records are already georeferenced?**_

----------

- Aggregators combined, accounting for obvious duplicates = `r format(records_georef, big.mark = ",")`
- Distinct localities represented by these records = `r nrow(localities_georef)`

----------

_**Of those records that are not already georeferenced, how many have something in the locality field?**_

----------

- Aggregators combined, accounting for obvious duplicates = `r format(nrow(records_locality), big.mark = ",")`
- Distinct localities represented by these records = `r nrow(localities_locality)`

----------

_**How many distinct collections are providing these specimen records?**_

----------

_`r nrow(providers_total)` data providers based on total number of records_

```{r echo = FALSE}
knitr::kable(providers_total) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "200px")
```

&nbsp;

_`r nrow(providers_ungeoreferenced)` data providers based on number of records needing to be georeferenced_

```{r echo = FALSE}
knitr::kable(providers_ungeoreferenced) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "200px")
```

----------

_**What is the geographic scope represented in these records, measured by country of collection?**_

----------

_`r nrow(countries_total)` values for country present in total number of records_

```{r echo = FALSE}
knitr::kable(countries_total) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "200px")
```

&nbsp;

_`r nrow(countries_ungeoreferenced)` values for country present in records needing to be georeferenced_

```{r echo = FALSE}
knitr::kable(countries_ungeoreferenced) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "200px")
```

----------

_**What is the temporal scope represented in these records, measured by year of collection?**_

----------

TBD

----------

_**What is the taxonomic scope represented in these records, measured by distinct species?**_

----------

_`r nrow(species_total)` values for species present in total number of records, `r format(round(sum(records_total_batnames/nrow(records)*100), 1), big.mark = ",")`% valid according to Batnames_

```{r echo = FALSE}
knitr::kable(species_total) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

&nbsp;

_`r nrow(species_ungeoreferenced)` values for species present in records needing to be georeferenced, `r format(round(sum(records_ungeoreferenced_batnames/nrow(records_locality)*100), 1), big.mark = ",")`% valid according to Batnames_

```{r echo = FALSE}
knitr::kable(species_ungeoreferenced) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

----------

_**What kind of ancillary locality information is represented in these records, measured by presence of values in the habitat field?**_

----------

- Total number of records with values in the habitat field = `r records_habitat` (`r round(sum(records_habitat/nrow(records)*100), 1)`%)

----------

_**What kind of ancillary locality information is represented in these records, measured by presence of values in the preparations and associatedSequences fields?**_

----------

- Total number of records with values in the preparations field = `r records_prep` (`r round(sum(records_prep/nrow(records)*100), 1)`%)
- Total number of records with values in the associatedSequences field = `r records_seq` (`r round(sum(records_seq/nrow(records)*100), 1)`%)
