---
title: "Exploratory Analysis for Georeferencing Project"
output:
  html_document:
    code_folding: hide
    df_print: kable
---

```{r message = FALSE}
# Load core libraries; install these packages if you have not already
library(ridigbio)
library(rgbif)
library(tidyverse)

# Load library for making nice HTML output
library(kableExtra)
```
```{r global_options, include = FALSE}
# Set global chunk options
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      results = 'asis')
```

----------

### Big picture

----------

```{r}
# Get count of mammal specimen records on iDigBio
mammalia_idb <- idig_count_records(rq = list(class = "mammalia"))

# Get count of mammal specimen records on GBIF
mammalia_gbif <- occ_count(taxonKey = 359,
                           basisOfRecord = "PRESERVED_SPECIMEN")

# Get count of mammal specimen records that are georeferenced on iDigBio
mammalia_georef_idb <- idig_count_records(rq = list(class = "mammalia",
                                                    geopoint = list(
                                                      type = "exists")))

# Get count of mammal specimen records that are georeferenced on GBIF
mammalia_georef_gbif <- occ_count(taxonKey = 359,
                                  basisOfRecord = "PRESERVED_SPECIMEN",
                                  georeferenced = TRUE)

# Get count of mammal specimens that are not georeferenced but do have values
# in the locality field on iDigBio
mammalia_locality_idb <- idig_count_records(rq = list(class = "mammalia",
                                                      geopoint = list(
                                                        type = "missing"),
                                                      locality = list(
                                                        type = "exists")))

# Get count of mammal specimens that are not georeferenced but do have values
# in the locality field on GBIF
mammalia_locality_gbif <- "TBD"
```

_**How many specimen records for Mammalia exist on iDigBio and GBIF?**_

- **iDigBio** = `r format(mammalia_idb, big.mark = ",")`
- **GBIF** = `r format(mammalia_gbif, big.mark = ",")`

_**How many of these specimen records are georeferenced already?**_

- **iDigBio** = `r format(mammalia_georef_idb, big.mark = ",")` (`r round(sum(mammalia_georef_idb/mammalia_idb*100), 1)`%)
- **GBIF** = `r format(mammalia_georef_gbif, big.mark = ",")` (`r round(sum(mammalia_georef_gbif/mammalia_gbif*100), 1)`%)

_**Of those specimen records that are not already georeferenced, how many have something in the locality field?**_

- **iDigBio** = `r format(mammalia_locality_idb, big.mark = ",")` (`r round(sum(mammalia_locality_idb/(mammalia_idb-mammalia_georef_idb)*100), 1)`%)
- **GBIF** = `r mammalia_locality_gbif`

----------

### Scoped by taxonomic priorities

----------

For each taxonomic scope, we want to know:

1. How many specimen records exist on iDigBio and on GBIF?
1. How many of these records are already georeferenced?
1. Of those records that are not already georeferenced, how many have something in the locality field?
1. How many distinct collections are providing these specimen records?
    - Who are the top five data providers based on total number of records?
    - What percentage of records are contributed by these top five?
    - Who are the top five data providers based on number of records needing to be georeferenced?
    - What percentage of records needing to be georeferenced are contributed by these top five?
1. Where are these recordsets being mobilized from? (e.g. the VertNet IPT)
1. What is the geographic scope represented in these records, measured by country of collection?
    - All records
    - Records in need of georeferencing
1. What is the temporal scope represented in these records, measured by year of collection?
    - All records
    - Records in need of georeferencing
1. What is the taxonomic scope represented in these records, measured by distinct species?
    - All records
    - Records in need of georeferencing
1. What kind of ancillary locality information is represented in these records, measured by presence of values in the habitat field?
    - All records
    - Records in need of georeferencing
1. What kind of ancillary locality information is represented in these records, measured by presence of values in the preparations and associatedSequences fields?
    - All records
    - Records in need of georeferencing

#### Primary taxonomic scope

RHINOLOPHIDAE, HIPPOSIDERIDAE, RHINONYCTERIDAE

```{r}
# Rhinonycteridae is a newly designated family and does not exist in the GBIF
# taxonomic backbone currently, so is not included in GBIF searches below.

# Get count of primary scope specimen records on iDigBio
primaryScope_idb <- idig_count_records(rq = list(family = c("Rhinolophidae", 
                                                            "Hipposideridae", 
                                                            "Rhinonycteridae")))

# Get count of primary scope specimen records on GBIF, by family
primaryScope_gbif_1 <- occ_count(taxonKey = 5449,
                           basisOfRecord = "PRESERVED_SPECIMEN")

# Get count of primary scope specimen records on GBIF, by family
primaryScope_gbif_2 <- occ_count(taxonKey = 3239753,
                           basisOfRecord = "PRESERVED_SPECIMEN")

# Sum family counts for primary scope specimen records on GBIF
primaryScope_gbif <- primaryScope_gbif_1 + primaryScope_gbif_2

# Get count of primary scope specimen records that are georeferenced on iDigBio
primaryScope_georef_idb <- idig_count_records(rq = list(family = c("Rhinolophidae", 
                                                            "Hipposideridae", 
                                                            "Rhinonycteridae"),
                                                    geopoint = list(
                                                      type = "exists")))

# Get count of primary scope specimen records that are georeferenced on GBIF
primaryScope_georef_gbif_1 <- occ_count(taxonKey = 5449,
                                  basisOfRecord = "PRESERVED_SPECIMEN",
                                  georeferenced = TRUE)

# Get count of primary scope specimen records that are georeferenced on GBIF, by family
primaryScope_georef_gbif_2 <- occ_count(taxonKey = 3239753,
                                  basisOfRecord = "PRESERVED_SPECIMEN",
                                  georeferenced = TRUE)

# Sum family counts for primary scope specimen records that are georeferenced on GBIF
primaryScope_georef_gbif <- primaryScope_georef_gbif_1 + primaryScope_georef_gbif_2

# Get count of primary scope specimens that are not georeferenced but do have values
# in the locality field on iDigBio
primaryScope_locality_idb <- idig_count_records(rq = list(family = c("Rhinolophidae", 
                                                            "Hipposideridae", 
                                                            "Rhinonycteridae"),
                                                      geopoint = list(
                                                        type = "missing"),
                                                      locality = list(
                                                        type = "exists")))

# Get count of mammal specimens that are not georeferenced but do have values
# in the locality field on GBIF
primaryScope_locality_gbif <- "TBD"
```

_**How many specimen records for the primary taxonomic scope exist on iDigBio and GBIF?**_

- **iDigBio** = `r format(primaryScope_idb, big.mark = ",")`
- **GBIF** = `r format(primaryScope_gbif, big.mark = ",")`

_**How many of these specimen records are georeferenced already?**_

- **iDigBio** = `r format(primaryScope_georef_idb, big.mark = ",")` (`r round(sum(primaryScope_georef_idb/primaryScope_idb*100), 1)`%)
- **GBIF** = `r format(primaryScope_georef_gbif, big.mark = ",")` (`r round(sum(primaryScope_georef_gbif/primaryScope_gbif*100), 1)`%)

_**Of those specimen records that are not already georeferenced, how many have something in the locality field?**_

- **iDigBio** = `r format(primaryScope_locality_idb, big.mark = ",")` (`r round(sum(primaryScope_locality_idb/(primaryScope_idb-primaryScope_georef_idb)*100), 1)`%)
- **GBIF** = 22,312 (51.6%)

```{r}
queryByFamily <- function(family, scope) {
  
  # Find GBIF taxonKey for given family
  taxonKey <- vector("integer", length(family))
  
  for (i in seq_along(family)) {
    taxonKey[[i]] <- name_backbone(name = family[[i]],
                            rank = "family",
                            class = "mammalia")
    }

  # Rename `taxonKey` with the input scope value and make it available
  # in the global environment
  # assign(paste("taxonKey_", scope, "Scope", sep = ""), taxonKey, envir = globalenv())
  
  # Get count of specimen records on iDigBio
  idb <- idig_count_records(rq = list(family = family))
  
  # Get count of specimen records that are georeferenced on iDigBio
  georef_idb <- idig_count_records(rq = list(family = family,
                                             geopoint = list(
                                               type = "exists")))
  
  # Get count of specimens that are not georeferenced but do have values
  # in the locality field on iDigBio
  locality_idb <- idig_count_records(rq = list(family = family,
                                               geopoint = list(
                                                 type = "missing"),
                                               locality = list(
                                                 type = "exists")))
  
  # Get count of specimen records on GBIF
  # gbif <- occ_count(taxonKey = taxonKey,
  #                  basisOfRecord = "PRESERVED_SPECIMEN")
  
  # Get count of specimen records that are georeferenced on GBIF
  # georef_gbif <- occ_count(taxonKey = taxonKey,
  #                         basisOfRecord = "PRESERVED_SPECIMEN",
  #                         georeferenced = TRUE)
  
  # Get count of specimens that are not georeferenced but do have values
  # in the locality field on GBIF
  locality_gbif <- "TBD"
  
  # Create table to present variables above
  summary <- tribble(
    ~source,
    ~totalCount,
    ~georefCount,
    ~georefPercent, 
    ~notGeorefCount,
    ~notGeorefPercent,
    ~couldGeorefCount,
    ~couldGeorefPercent,
    "iDigBio",
    idb,
    georef_idb,
    paste(round(sum(georef_idb/idb*100), 1), "%", sep = ""),
    sum(idb-georef_idb),
    paste(round(sum((idb-georef_idb)/idb*100), 1), "%", sep = ""),
    locality_idb,
    paste(round(sum(locality_idb/(idb-georef_idb)*100), 1), "%", sep = ""),
    "GBIF",
    gbif,
    georef_gbif,
    paste(round(sum(georef_gbif/gbif*100), 1), "%", sep = ""),
    sum(gbif-georef_gbif),
    paste(round(sum((gbif-georef_gbif)/gbif*100), 1), "%", sep = ""),
    locality_gbif,
    "TBD")
    
  # Rename `summary` with the input scope value and make it available
  # in the global environment
  assign(paste("summary_", scope, "Scope", sep = ""), summary, envir = globalenv())
  
  # Pull in specimen records from iDigBio that are not georeferenced but that do
  # have values in the locality field
  records_idb <- idig_search_records(rq = list(family = family,
                                               geopoint = list(
                                                 type = "missing"),
                                               locality = list(
                                                 type = "exists")),
                                     fields = c("uuid",
                                                "recordset",
                                                "institutioncode",
                                                "collectioncode",
                                                "catalognumber",
                                                "scientificname",
                                                "eventdate",
                                                "country",
                                                "stateprovince",
                                                "county",
                                                "municipality",
                                                "data.dwc:locality",
                                                "data.dwc:habitat",
                                                "data.dwc:preparations",
                                                "data.dwc:associatedSequences"))
  
  # Create column in `records_idb` to match duplicate records with GBIF data
  records_idb <- records_idb %>% 
    unite(matchDuplicates, 
          c("institutioncode", "collectioncode", "catalognumber"), 
          sep = " ", 
          remove = FALSE) %>% 
    mutate(matchDuplicates = tolower(matchDuplicates))
  
  # Pull in specimen records from GBIF that are not georeferenced
  records_gbif <- occ_search(taxonKey = taxonKey,
                             basisOfRecord = "PRESERVED_SPECIMEN",
                             hasCoordinate = FALSE,
                             fields = c("key",
                                        "datasetKey",
                                        "institutionCode",
                                        "collectionCode",
                                        "catalogNumber",
                                        "acceptedScientificName",
                                        "year",
                                        "country",
                                        "stateProvince",
                                        "county",
                                        "municipality",
                                        "locality",
                                        "habitat",
                                        "preparations",
                                        "associatedSequences"),
                             limit = 100000)
  
  # Extract dataframe from `records_gbif` list
  records_gbif_1 <- records_gbif$`5449`$data
  records_gbif_2 <- records_gbif$`3239753`$data
  records_gbif <- bind_rows(list(records_gbif_1, records_gbif_2), .id = "id")
  
  # Subset `records_gbif` to include only those records with locality values,
  # and create column to match duplicate records with iDigBio data
  records_gbif <- records_gbif %>% 
    filter(!is.na(locality)) %>% 
    unite(matchDuplicates, 
          c("institutionCode", "collectionCode", "catalogNumber"), 
          sep = " ", 
          remove = FALSE) %>% 
    mutate(matchDuplicates = tolower(matchDuplicates))
  
  # Join `records_idb` and `records_gbif`
  records <- records_gbif %>% 
    full_join(records_idb, by = "matchDuplicates") %>% 
    mutate(idigbio = case_when(!is.na(uuid) ~ "yes")) %>% 
    mutate(gbif = case_when(!is.na(key) ~ "yes")) %>% 
    mutate(institutionCode = tolower(coalesce(institutioncode, institutionCode))) %>% 
    mutate(collectionCode = tolower(coalesce(collectioncode, collectionCode))) %>% 
    mutate(catalogNumber = tolower(coalesce(catalognumber, catalogNumber))) %>% 
    mutate(country = tolower(coalesce(country.x, country.y))) %>% 
    mutate(stateProvince = tolower(coalesce(stateprovince, stateProvince))) %>% 
    mutate(county = tolower(coalesce(county.x, county.y))) %>% 
    mutate(municipality = tolower(coalesce(municipality.x, municipality.y))) %>% 
    mutate(locality = coalesce(`data.dwc:locality`, locality)) %>% 
    mutate(habitat = coalesce(`data.dwc:habitat`, habitat)) %>% 
    mutate(preparations = coalesce(`data.dwc:preparations`, preparations)) %>% 
    mutate(year_idb = as.integer(format(lubridate::as_date(eventdate), "%Y"))) %>% 
    mutate(year = coalesce(year, year_idb)) %>% 
    select(idigbio,
           gbif,
           idigbio_uuid = uuid, 
           gbif_key = key, 
           idigbio_recordset = recordset, 
           gbif_datasetkey = datasetKey, 
           institutionCode,
           collectionCode,
           catalogNumber,
           idigbio_scientificname = scientificname,
           gbif_acceptedScientificName = acceptedScientificName,
           country,
           stateProvince,
           county,
           municipality,
           locality,
           year,
           habitat,
           preparations,
           idigbio_associatedSequences = `data.dwc:associatedSequences`) %>% 
    arrange(country, stateProvince, municipality)

  # Rename `records` with the input family value and make it available
  # in the global environment
  assign(paste("records_", scope, "Scope", sep = ""), records, envir = globalenv())
  
  # Assign input family value to filename for outputing `records`
  filename <- paste("records_", scope, "Scope", ".csv", sep = "")
  
  # Output `records` to a CSV file
  write_csv(records, filename)
  
  # Create table to present summary data from `records`
  summaryCombo <- tribble(
    ~totalCount,
    ~totalCount_idigbio,
    ~totalCount_gbif,
    ~uniqueInstitutions,
    ~uniqueCountries,
    ~uniqueLocalities,
    ~percentRows_habitat,
    ~percentRows_prep,
    ~percentRows_sequence,
    nrow(records),
    nrow(filter(records, idigbio=="yes")),
    nrow(filter(records, gbif=="yes")),
    nrow(tally(group_by(records, institutionCode))),
    nrow(tally(group_by(records, country))),
    nrow(tally(group_by(records, locality))),
    round(sum(nrow(filter(records, !is.na(habitat)))/nrow(records)*100), 1),
    round(sum(nrow(filter(records, !is.na(preparations)))/nrow(records)*100), 1),
    round(sum(nrow(filter(records, !is.na(idigbio_associatedSequences)))/nrow(records)*100), 1))
    
  # Rename `summaryCombo` with the input family value and make it available
  # in the global environment
  assign(paste("summaryCombo_", scope, "Scope", sep = ""), summaryCombo, envir = globalenv())
  
}

# queryByFamily(family = c("Rhinolophidae", "Hipposideridae", "Rhinonycteridae"),
#              scope = "primary")
```
```{r}
# Save time rendering RMarkdown by reading in data from previously-run queries
summaryCombo_primaryScope <- read_csv("summaryCombo_primaryScope.csv")
records_primaryScope <- read_csv("records_primaryScope.csv")
```
##### Data that is not already georeferenced but has locality information

Summaries below are from iDigBio and GBIF data combined to have a single record per unique combination of institution code + collection code + catalog number.

_Summary of combined data_

```{r echo = FALSE}
knitr::kable(summaryCombo_primaryScope) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%")
```

_Top data providers _

```{r echo = FALSE, results = 'asis'}
records_primaryScope %>% 
  group_by(institutionCode) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  mutate(recordPercent = round(n/sum(n)*100, 1)) %>% 
  rename(recordCount = n) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "300px")
```

_Unique values for country_

```{r echo = FALSE, results = 'asis'}
records_primaryScope %>% 
  group_by(country) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  mutate(recordPercent = round(n/sum(n)*100, 1)) %>% 
  rename(recordCount = n) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "300px")
```

#### Secondary taxonomic scope

MEGADERMATIDAE, RHINOPOMATIDAE, CRASEONYCTERIDAE

#### Tertiary taxonomic scope

PTEROPODIDAE

#### Broadest taxonomic scope

Pangolins: MANIDAE, PATRIOMANIDAE, ESCAVADODONTIDAE

Civets:

- Arctogalidia trivirgata
- Macrogalidia musschenbroekii
- Paguma larvata
- Paradoxurus hermaphroditus
- Paradoxurus jerdoni
- Paradoxurus zeylonensis
- Chrotogale owstoni
- Cynogale bennettii
- Diplogale hosei
- Hemigalus derbyanus
- Civettictis civetta
- Viverra civettina
- Viverra megaspila
- Viverra tangalunga
- Viverra zibetha
- Viverricula indica
- Fossa fossana
- Nandinia binotata
